Advancing Sustainable Mobility: A GIS-Based Multi-Criteria Decision Analysis for Equitable Electric Vehicle Supply Equipment Deployment in Philadelphia

### 2025-02-18 code/MCDA-EVSE.Rmd
fix function

@@

@@

## cluster 1
overall_c1 <- TOPSIS(performanceTable_c1, weights_c1, criteriaMinMax)
TOPSIS_c1 <- data.frame(uniqueID = names(overall_c1), values = overall_c1) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c1 = rank(values),
         topsis_percentile_c1 = percent_rank(topsis_rank_c1) * 100)

# test for original
overall_c1_org <- TOPSIS(performanceTable_c1, weights_original, criteriaMinMax)
TOPSIS_c1_org <- data.frame(uniqueID = names(overall_c1_org), values = overall_c1_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c1_org = rank(values),
         topsis_percentile_c1_org = percent_rank(topsis_rank_c1_org) * 100)

## cluster 2
overall_c2 <- TOPSIS(performanceTable_c2, weights_c2, criteriaMinMax)
TOPSIS_c2 <- data.frame(uniqueID = names(overall_c2), values = overall_c2) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c2 = rank(values),
         topsis_percentile_c2 = percent_rank(topsis_rank_c2) * 100)

# test for original
overall_c2_org <- TOPSIS(performanceTable_c2, weights_original, criteriaMinMax)
TOPSIS_c2_org <- data.frame(uniqueID = names(overall_c2_org), values = overall_c2_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c2_org = rank(values),
         topsis_percentile_c2_org = percent_rank(topsis_rank_c2_org) * 100)

## cluster 3
overall_c3 <- TOPSIS(performanceTable_c3, weights_c3, criteriaMinMax)
TOPSIS_c3 <- data.frame(uniqueID = names(overall_c3), values = overall_c3) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c3 = rank(values),
         topsis_percentile_c3 = percent_rank(topsis_rank_c3) * 100)

# test for original
overall_c3_org <- TOPSIS(performanceTable_c3, weights_original, criteriaMinMax)
TOPSIS_c3_org <- data.frame(uniqueID = names(overall_c3_org), values = overall_c3_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c3_org = rank(values),
         topsis_percentile_c3_org = percent_rank(topsis_rank_c3_org) * 100)

## cluster 4
overall_c4 <- TOPSIS(performanceTable_c4, weights_c4, criteriaMinMax)
TOPSIS_c4 <- data.frame(uniqueID = names(overall_c4), values = overall_c4) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c4 = rank(values),
         topsis_percentile_c4 = percent_rank(topsis_rank_c4) * 100)

# test for original
overall_c4_org <- TOPSIS(performanceTable_c4, weights_original, criteriaMinMax)
TOPSIS_c4_org <- data.frame(uniqueID = names(overall_c4_org), values = overall_c4_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c4_org = rank(values),
         topsis_percentile_c4_org = percent_rank(topsis_rank_c4_org) * 100)

We ran the TOPSIS MCDA method. We computed the local rank as well as rank percentile for each cluster using local weights and the original global weights respectively. We compared that to the percentile of global rank using TOPSIS as well as the average global rank percentile.

Comparing local vs. global TOPSIS reveals how rankings shift due to local conditions.
Comparing local TOPSIS to global average MCDA checks if local patterns align with broader decision-making trends.
Comparing local weights vs. global weights in local TOPSIS isolates the effect of weight adjustments on rankings.

# Get the original rank and compute percentile rank 
# then join it with the result from local MCDA
# to compare whether the percentile rank is higher or lower than the global MCDA
final_percentile <- final %>% 
  dplyr::select(NAME, uniqueID, avg_rank, topsis_rank, wsm_rank) %>%
  mutate(
    avg_percentile = percent_rank(avg_rank) * 100,
    topsis_percentile = percent_rank(topsis_rank) * 100)

@@

# quick check for cluster 1
TOPSIS_c1  <- TOPSIS_c1 %>% 
  left_join(TOPSIS_c1_org %>% dplyr::select(uniqueID, topsis_percentile_c1_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile,

@@

# quick check for cluster 2
TOPSIS_c2  <- TOPSIS_c2 %>% 
  left_join(TOPSIS_c2_org %>% dplyr::select(uniqueID, topsis_percentile_c2_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, topsis_percentile, topsis_percentile_c2, topsis_percentile_c2_org)

# quick check for cluster 3
TOPSIS_c3  <- TOPSIS_c3 %>% 
  left_join(TOPSIS_c3_org %>% dplyr::select(uniqueID, topsis_percentile_c3_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, topsis_percentile, topsis_percentile_c3, topsis_percentile_c3_org)

# quick check for cluster 4
TOPSIS_c4  <- TOPSIS_c4 %>% 
  left_join(TOPSIS_c4_org %>% dplyr::select(uniqueID, topsis_percentile_c4_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, topsis_percentile, topsis_percentile_c4, topsis_percentile_c4_org)

Now do some visualization and analysis for comparison using cluster 4 as an example.

# using cluster 1 as an example
TOPSIS_c4 %>% 
  ggplot(aes(x = topsis_percentile_c4_org, y = topsis_percentile_c4)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) + 
  labs(title = "Comparison of Global and Local TOPSIS Percentile Rank for Cluster 1",
       x = "Global Percentile Rank",
       y = "Local Percentile Rank") +
  theme_minimal() +
  theme(plot.subtitle = element_text(size = 12,face = "italic"),
        plot.title = element_text(size = 15, face = "bold"), 
        axis.text.x=element_text(size=8),
        axis.text.y=element_text(size=8), 
        axis.title=element_text(size=9), 
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth =0.8))

model1 <- lm(topsis_percentile_c3 ~ topsis_percentile, data = TOPSIS_c3)
model2<- lm(topsis_percentile_c3 ~ topsis_percentile_c3_org, data = TOPSIS_c3)
model3 <- lm(topsis_percentile ~ topsis_percentile_c3_org, data = TOPSIS_c3)

Cluster 1: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- -0.870
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.901
Local TOPSIS (Global Weights) vs. Global TOPSIS --- -0.207

Cluster 2: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- -0.979
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.930
Local TOPSIS (Global Weights) vs. Global TOPSIS --- -0.802

Cluster 3:
Local TOPSIS (Local Weights) vs. Global TOPSIS --- -0.853
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.911
Local TOPSIS (Global Weights) vs. Global TOPSIS --- -0.890

Cluster 4: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- -0.837
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.962
Local TOPSIS (Global Weights) vs. Global TOPSIS --- -0.491

# an overall regression summary table here

@@

TOPSIS_c4 %>% 
  left_join(cluster4 %>% dplyr::select(uniqueID, geometry), by = "uniqueID") %>% 
  st_as_sf() %>% 
  ggplot() +
  geom_sf(data = fishnet, color = "black") +
  geom_sf(aes(fill = topsis_percentile_c4), color = "black") +
  scale_fill_viridis_c(option = "plasma") +
  theme_bw()

>>>>

We will perform three sets of comparison. 
- Global MCDA (Global Weights) ranks grids based on criteria importance at the citywide scale. This approach ensures consistency across Philadelphia but may overlook local variations.
- Local MCDA (Local Weights) adapts weights to the unique characteristics of each cluster. This method highlights localized priorities that might be underrepresented in a citywide analysis.
- Local MCDA (Global Weights) applies the citywide weight structure within each cluster, allowing us to observe how the same criteria influence different local contexts.

@@

# Get the original rank and compute percentile rank 
# then join it with the result from local MCDA
# to compare whether the percentile rank is higher or lower than the global MCDA
final_percentile <- final %>% 
  dplyr::select(NAME, uniqueID, avg_rank, topsis_rank, wsm_rank) %>%
  mutate(
    avg_percentile = percent_rank(avg_rank) * 100,
    wsm_percentile = percent_rank(wsm_rank) * 100,
    topsis_percentile = percent_rank(topsis_rank) * 100)

## Try Local WSM

```{r}
# weights_c1 <- c(0.12, 0.14, 0.12, 0.16, 0.16, 0.08, 0.06, 0.06, 0.10)
WSM_c1 <- cluster1 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.12*scl_evse_re + 0.14*scl_density + 0.12*scl_drivingpop + 0.16*scl_acc_re + 0.16*scl_elec + 0.08*scl_parking_re + 0.06*scl_policefire_re + 0.06*zone_score + 0.10*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>% 
  mutate(wsm_rank_c1 = min_rank(desc(score))) %>% 
  mutate(wsm_percentile_c1 = percent_rank(wsm_rank_c1) * 100)

#weights_c2 <- c(0.14, 0.08, 0.08, 0.10, 0.12, 0.12, 0.06, 0.10, 0.20)
WSM_c2 <- cluster2 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% # now inverse some scales
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.14*scl_evse_re + 0.08*scl_density + 0.08*scl_drivingpop + 0.10*scl_acc_re + 0.12*scl_elec + 0.12*scl_parking_re + 0.06*scl_policefire_re + 0.10*zone_score + 0.20*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c2 = min_rank(desc(score))) %>%
  mutate(wsm_percentile_c2 = percent_rank(wsm_rank_c2) * 100)

#weights_c3 <- c(0.20, 0.10, 0.12, 0.14, 0.18, 0.06, 0.06, 0.06, 0.08)
WSM_c3 <- cluster3 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.20*scl_evse_re + 0.10*scl_density + 0.12*scl_drivingpop + 0.14*scl_acc_re + 0.18*scl_elec + 0.06*scl_parking_re + 0.06*scl_policefire_re + 0.06*zone_score + 0.08*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c3 = min_rank(desc(score))) %>%
  mutate(wsm_percentile_c3 = percent_rank(wsm_rank_c3) * 100)

#weights_c4 <- c(0.16, 0.14, 0.12, 0.18, 0.12, 0.08, 0.04, 0.06, 0.10)
WSM_c4 <- cluster4 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.16*scl_evse_re + 0.14*scl_density + 0.12*scl_drivingpop + 0.18*scl_acc_re + 0.12*scl_elec + 0.08*scl_parking_re + 0.04*scl_policefire_re + 0.06*zone_score + 0.10*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c4 = min_rank(desc(score))) %>%
  mutate(wsm_percentile_c4 = percent_rank(wsm_rank_c4) * 100)

```{r test with original}

# using original weights
# this is evse.nn, density, driving_pop, TSFCA, elec_cnt, parking.nn, policefire.nn, zone score, elect_use
# weights_original <- c(0.18, 0.12, 0.1, 0.12, 0.15, 0.1, 0.05, 0.05, 0.13)

WSM_c1_org <- cluster1 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>% 
  mutate(wsm_rank_c1_org = min_rank(desc(score))) %>% 
  mutate(wsm_percentile_c1_org = percent_rank(wsm_rank_c1_org) * 100)

WSM_c2_org <- cluster2 %>%
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c2_org = min_rank(desc(score))) %>%
  mutate(wsm_percentile_c2_org = percent_rank(wsm_rank_c2_org) * 100)

WSM_c3_org <- cluster3 %>%
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c3_org = min_rank(desc(score))) %>%
  mutate(wsm_percentile_c3_org = percent_rank(wsm_rank_c3_org) * 100)

WSM_c4_org <- cluster4 %>%
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c4_org = min_rank(desc(score))) %>%
  mutate(wsm_percentile_c4_org = percent_rank(wsm_rank_c4_org) * 100)

# join with original percentile
WSM_c1_cp <- WSM_c1 %>% 
  st_drop_geometry() %>%
  left_join(WSM_c1_org %>% dplyr::select(uniqueID, wsm_percentile_c1_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c1, wsm_percentile_c1_org)

WSM_c2_cp <- WSM_c2 %>%
  st_drop_geometry() %>%
  left_join(WSM_c2_org %>% dplyr::select(uniqueID, wsm_percentile_c2_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,

                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c2, wsm_percentile_c2_org)

WSM_c3_cp <- WSM_c3 %>%
  st_drop_geometry() %>%
  left_join(WSM_c3_org %>% dplyr::select(uniqueID, wsm_percentile_c3_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,

                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c3, wsm_percentile_c3_org)

WSM_c4_cp <- WSM_c4 %>%
  st_drop_geometry() %>%
  left_join(WSM_c4_org %>% dplyr::select(uniqueID, wsm_percentile_c4_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,

                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c4, wsm_percentile_c4_org)

WSM_c3_cp %>% 
  ggplot(aes(x = wsm_percentile , y = wsm_percentile_c3)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) + 
  labs(title = "Comparison of Global and Local TOPSIS Percentile Rank for Cluster 1",
       x = "Global Percentile Rank",
       y = "Local Percentile Rank") +
  theme_minimal() +
  theme(plot.subtitle = element_text(size = 12,face = "italic"),
        plot.title = element_text(size = 15, face = "bold"), 
        axis.text.x=element_text(size=8),
        axis.text.y=element_text(size=8), 
        axis.title=element_text(size=9), 
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth =0.8))

model1_wsm <- lm(wsm_percentile_c4 ~ wsm_percentile, data = WSM_c4_cp)
model2_wsm<- lm(wsm_percentile_c4 ~ wsm_percentile_c4_org, data = WSM_c4_cp)
model3_wsm <- lm(wsm_percentile_c4_org ~ wsm_percentile , data = WSM_c4_cp)

library(stargazer)
stargazer(model1_wsm, model2_wsm, model3_wsm, type = "text")
```

Cluster 1: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 2.800 (0.787 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.973 (0.947 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.269 (0.724 Rsq)

Cluster 2: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 1.240 (0.795 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.987 (0.975 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.654 (0.821 Rsq)

Cluster 3:
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 1.060 (0.904 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.982 (0.965 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.841 (0.879 Rsq)

Cluster 4: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 1.220 (0.725 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.977 (0.954 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.609 (0.76 Rsq)

High R² (0.947–0.975) for Local MCDA (Global Weights) vs. Local MCDA (Local Weights) across all clusters suggests that applying local weights slightly modifies but does not drastically change rankings. 
Lower slopes and R² for Local MCDA (Local Weights) vs. Global MCDA (especially in Cluster 1 and Cluster 4) indicate that the most significant rank shifts occur when transitioning from a global model to a local model.

In cluster 1, the slope of 2.800 for Local (Local Weights) vs. Global MCDA suggests extreme shifts in ranking, with certain grids ranking much higher or much lower locally than globally. Locally, some areas within the core have high demand but remain underserved, which global analysis fails to highlight.

In cluster 2, the slope for Local (Local Weights) vs. Global MCDA is 1.240 with R² of 0.795, suggesting that while local factors alter rankings, the overall prioritization is more aligned with the global MCDA compared to Cluster 1.

In cluster 3, the slope of 1.060 for Local (Local Weights) vs. Global MCDA, with an R² of 0.904, indicates minimal ranking disruptions between global and local approaches.

In cluster 4, the slope for Local (Local Weights) vs. Global MCDA is 1.220 with an R² of 0.725, suggesting substantial local ranking shifts. Additionally, the slope of 0.609 for Global MCDA vs. Local MCDA (Global Weights) is among the lowest, indicating that local constraints heavily influence rankings.

The choice of local vs. global scale significantly alters rankings, particularly in areas with high spatial and socioeconomic variability (e.g., Cluster 1 and Cluster 4).Localized MCDA is crucial for equity-driven EVSE planning, as global approaches may overlook accessibility barriers in disadvantaged communities. Areas with high demand but infrastructure gaps (e.g., Cluster 1 and Cluster 4) require localized adjustments.

See an example here: 
842 Frankford. in Cluster 4
Ranked 45 in Global
Ranked 16 in Local

## Try Local TOPSIS

@@

## cluster 1
overall_c1 <- TOPSIS(performanceTable_c1, weights_c1, criteriaMinMax)
TOPSIS_c1 <- data.frame(uniqueID = names(overall_c1), values = overall_c1) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c1 = min_rank(desc(values)),
         topsis_percentile_c1 = percent_rank(topsis_rank_c1) * 100)

# test for original
overall_c1_org <- TOPSIS(performanceTable_c1, weights_original, criteriaMinMax)
TOPSIS_c1_org <- data.frame(uniqueID = names(overall_c1_org), values = overall_c1_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c1_org = min_rank(desc(values)),
         topsis_percentile_c1_org = percent_rank(topsis_rank_c1_org) * 100)

## cluster 2
overall_c2 <- TOPSIS(performanceTable_c2, weights_c2, criteriaMinMax)
TOPSIS_c2 <- data.frame(uniqueID = names(overall_c2), values = overall_c2) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c2 = min_rank(desc(values)),
         topsis_percentile_c2 = percent_rank(topsis_rank_c2) * 100)

# test for original
overall_c2_org <- TOPSIS(performanceTable_c2, weights_original, criteriaMinMax)
TOPSIS_c2_org <- data.frame(uniqueID = names(overall_c2_org), values = overall_c2_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c2_org = min_rank(desc(values)),
         topsis_percentile_c2_org = percent_rank(topsis_rank_c2_org) * 100)

## cluster 3
overall_c3 <- TOPSIS(performanceTable_c3, weights_c3, criteriaMinMax)
TOPSIS_c3 <- data.frame(uniqueID = names(overall_c3), values = overall_c3) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c3 = min_rank(desc(values)),
         topsis_percentile_c3 = percent_rank(topsis_rank_c3) * 100)

# test for original
overall_c3_org <- TOPSIS(performanceTable_c3, weights_original, criteriaMinMax)
TOPSIS_c3_org <- data.frame(uniqueID = names(overall_c3_org), values = overall_c3_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c3_org = min_rank(desc(values)),
         topsis_percentile_c3_org = percent_rank(topsis_rank_c3_org) * 100)

## cluster 4
overall_c4 <- TOPSIS(performanceTable_c4, weights_c4, criteriaMinMax)
TOPSIS_c4 <- data.frame(uniqueID = names(overall_c4), values = overall_c4) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c4 = min_rank(desc(values)),
         topsis_percentile_c4 = percent_rank(topsis_rank_c4) * 100)

# test for original
overall_c4_org <- TOPSIS(performanceTable_c4, weights_original, criteriaMinMax)
TOPSIS_c4_org <- data.frame(uniqueID = names(overall_c4_org), values = overall_c4_org) %>%  
  mutate(uniqueID = as.integer(uniqueID)) %>% 
  mutate(topsis_rank_c4_org = min_rank(desc(values)),
         topsis_percentile_c4_org = percent_rank(topsis_rank_c4_org) * 100)

@@

# quick check for cluster 1
TOPSIS_c1_cp  <- TOPSIS_c1 %>% 
  left_join(TOPSIS_c1_org %>% dplyr::select(uniqueID, topsis_percentile_c1_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile,

@@

# quick check for cluster 2
TOPSIS_c2_cp  <- TOPSIS_c2 %>% 
  left_join(TOPSIS_c2_org %>% dplyr::select(uniqueID, topsis_percentile_c2_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, topsis_percentile, topsis_percentile_c2, topsis_percentile_c2_org)

# quick check for cluster 3
TOPSIS_c3_cp  <- TOPSIS_c3 %>% 
  left_join(TOPSIS_c3_org %>% dplyr::select(uniqueID, topsis_percentile_c3_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, topsis_percentile, topsis_percentile_c3, topsis_percentile_c3_org)

# quick check for cluster 4
TOPSIS_c4_cp  <- TOPSIS_c4 %>% 
  left_join(TOPSIS_c4_org %>% dplyr::select(uniqueID, topsis_percentile_c4_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               topsis_percentile, 
                                               avg_percentile
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, topsis_percentile, topsis_percentile_c4, topsis_percentile_c4_org)

@@

# using cluster 1 as an example
TOPSIS_c1_cp %>% 
  ggplot(aes(x = topsis_percentile_c1_org, y = topsis_percentile)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  geom_smooth(method = "lm", color = "blue", se = FALSE) + 
  labs(title = "Comparison of Global and Local TOPSIS Percentile Rank for Cluster 1",
       x = "Global Percentile Rank",
       y = "Local Percentile Rank") +
  theme_minimal() +
  theme(plot.subtitle = element_text(size = 12,face = "italic"),
        plot.title = element_text(size = 15, face = "bold"), 
        axis.text.x=element_text(size=8),
        axis.text.y=element_text(size=8), 
        axis.title=element_text(size=9), 
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth =0.8))

model1 <- lm(topsis_percentile_c1 ~ topsis_percentile, data = TOPSIS_c1_cp)
model2<- lm(topsis_percentile_c1 ~ topsis_percentile_c1_org, data = TOPSIS_c1_cp)
model3 <- lm(topsis_percentile ~ topsis_percentile_c1_org, data = TOPSIS_c1_cp)

Cluster 1: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 0.870 (0.305 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.901 (0.811 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.207 (0.106 Rsq)

Cluster 2: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 0.979 (0.868 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.930 (0.865 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.802 (0.711 Rsq)

Cluster 3:
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 0.853 (0.727 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.911 (0.830 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.890 (0.794 Rsq)

Cluster 4: 
Local TOPSIS (Local Weights) vs. Global TOPSIS --- 0.837 (0.416 Rsq)
Local TOPSIS (Global Weights) vs. Local TOPSIS (Local Weights) --- 0.962 (0.925 Rsq)
Local TOPSIS (Global Weights) vs. Global TOPSIS --- 0.491 (0.406)

This analysis explores the relationship between local and global TOPSIS approaches across four clusters, evaluating how different weighting strategies impact decision-making outcomes. Correlation coefficients (r) and R-squared (R²) values provide insights into the alignment between methods and the extent to which local preferences (either local or global weights) deviate from global decision-making frameworks.
